Unveiling Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects

### 2014-09-04 android-stacks.tex
results - question 1

Besides, the runtime exceptions are also implicitly thrown by the JVM when a program violates 
the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
error, null pointer references).\footnote{Some programming languages react to such errors by peremptorily terminating the program; 
other programming languages allow similar approaches e.g.C\#; and others let the program to continue
 its execution in some situations such as the out-of-bounds array index e.g., C++. }  
The programs are not expected to handle such JVM runtime exceptions ~\cite{gosling2000java} s.

@@

In Java, once an exception is thrown, the JVM looks for the nearest enclosing exception handler
(in Java it is represented by the try-catch block), and unwinds the execution stack if necessary.
 Looking for the handler on the invocation stack is claimed to increase software reusability, 
since the invoker of an operation can handle it in a wider context ~/cite{miller1997issues}.

@@

\emph{II-Error represents an unrecoverable condition detected by the JVM which
should not he handled} (\cite{gosling2000java}). It results from failures detected
by the Java Virtual Machine which indicate resource deficiencies, invariant
failures or other conditions that make impossible the program to recover.

@@

\emph{III-A method should throw exceptions that precisely define the
exceptional condition} (\cite{gosling2000java,bloch2008effective}). To do so,
developers should either try to reuse the exception types already defined in the
JVM or they should create a specific exception. Throwing general types such as a
pure java.lang.Exception or a java.lang.RuntimeException is considered a bad practice.

@@

\emph{IV- All exceptions explicitly thrown by a method should be documented.}
(\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled JVM runtime exceptions related to programming
mistakes} as carefully as checked exceptions. Doing so, the clients of a method
will be aware of all the exceptions the method can throw and can protect design the code
to deal with them ~\cite{Robil00}.

@@

\subsection{Manual Inspections}
As mentioned before, this mixed-methods study was also supported by manual inspections. 
Manual inspections were used, in the context of this study to support: (i) the identification of Android projects (described previously);
(ii) and to identify the type of some exceptions reported on issues, that could not be automatically identified by the ExceptionMiner tool.
Some exceptions could not be automaticaly identified because they were 
not defined on the last version of the application (the one analyzed by ExceptionMiner),
or were defined on libraries not found among application resources.
In such cases we manually inspect the sourcecode or javadoc
 documentation available online of each of them. 
When the exception was not found on the previous versions, the exception where classified 
in our study as "Undefined".  Only 31 exceptions 
remained undefined - such exceptions were reported on 61 exception stack traces as 
presented in Table ~\ref{tab:typeroottab}.

@@

@@

This section presents the results of the study, providing both a
quantitative and qualitative analysis of the outcomes. The structure of this section
 is centered on questions that guied the search for 
source-code fragilities based on exception stack trace information.

@@

After distilling the information available on the exception stack traces we could find 
the the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:topten} illustrates the top 10 root causes found in the study,
 ranked by the number of distinct project repositories on which they were reported. 
The table shows how many times method responsible for signaling the exceptions were
part of Android Platform, the application itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers}

% \footnote{Nine exception stack traces contained only elements of JVM and were discarded from the analysis}.

%For each exception we could discover the number of times it was reported as the root caused on exception stack traces,
%the number of distinct projects on which such stacks were defined (and consequently its
%popularity) and the of times it was signaled by different signaler types. We
%could calculate the popularity of each root cause, considering the number of
%distinct projects they were found and the whole set of projects analyzed with at
%least one stack trace (482). Table~\ref{tab:topten} 
%and Figure~\ref{fig:androidsignaler} presents the mined data.

%SELECTS:
%SELECT COUNT(distinct(repo)) FROM STACKTRACESUMMARY WHERE MAIN_CAUSE LIKE '%java.lang.NullPointerException%'
%and repo !='android';

%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE MAIN_CAUSE LIKE '%java.lang.IllegalStateException%'
%and repo !='android' AND MAINSIGNALER_FROM LIKE '%ANDROID%';

%java.lang.NullPointerException            & 254 /  78  & 52,70\%  & 1225 / 439  & 30,08\% & 473 /  26  & 18 /  & 595 / 241 & 137 / 169 \\  %LIBORAPP=408
%java.lang.IllegalStateException          & 99 / 21  & 20,54\%  & 234 / 44  & 5,75\%  & 165 / 15 & 12  / 0 & 36  / 5 & 20 / 24  \\
%java.lang.IllegalArgumentException          & 94 / 48  & 19,50\% & 255 / 98  & 6,26\% & 146 / 41 & 6 / 0 & 64 / 31  & 39  / 26  \\
%java.lang.RuntimeException                & 91 / 31 & 18,88\% & 232 / 87  & 5,70\%  & 167 / 33 & 1 / 0 & 47 / 17 & 17 / 37   \\
%java.lang.OutOfMemoryError                 & 56  / 22  & 11,62\% & 180 / 57 & 4,42\% & 121 / 18 & 15 / 0 & 17 / 18  & 23 / 13  \\
%java.lang.NoClassDefFoundError           & 52  / 15 & 10,79\%  & 73 / 21   & 1,79\%  & 9 / 1   & 0 / 0  & 37 / 9  & 26 / 11   \\
%java.lang.ClassCastException             & 49  / 15 & 10,17\%  & 94 / 36   & 2,31\%  & 43 / 3  & 0  / 0 & 40 / 15  & 11 / 18  \\
%java.lang.IndexOutOfBoundsException       & 47  / 15 & 9,75\%   & 127 / 39  & 3,12\% & 47 / 2  & 0/0  & 71/22  & 8 / 14    \\
%java.lang.NoSuchMethodError               & 40 / 14  & 8,30\%  & 57 / 23  & 1,40\%  & 9  /  0  & 0  / 0 & 39 /  17    & 9 / 6  \\
%java.util.ConcurrentModificationException & 38 / 5 & 7,88\% & 54 / 11  & 1,33\%   & 5  / 0  & 0  / 0 & 43 / 3 & 6 / 7  \\

%java.lang.IllegalStateException          & 99 / 21  & 20,54\%  & 234 / 44  & 5,75\%  & 165 / 15 & 12  / 0 & 36  / 5 & 20 / 24  \\
%java.lang.IllegalArgumentException          & 94 / 48  & 19,50\% & 255 / 98  & 6,26\% & 146 / 41 & 6 / 0 & 64 / 31  & 39  / 26  \\
%java.lang.RuntimeException                & 91 / 31 & 18,88\% & 232 / 87  & 5,70\%  & 167 / 33 & 1 / 0 & 47 / 17 & 17 / 37   \\
%java.lang.OutOfMemoryError                 & 56  / 22  & 11,62\% & 180 / 57 & 4,42\% & 121 / 18 & 15 / 0 & 17 / 18  & 23 / 13  \\
%java.lang.NoClassDefFoundError           & 52  / 15 & 10,79\%  & 73 / 21   & 1,79\%  & 9 / 1   & 0 / 0  & 37 / 9  & 26 / 11   \\
%java.lang.ClassCastException             & 49  / 15 & 10,17\%  & 94 / 36   & 2,31\%  & 43 / 3  & 0  / 0 & 40 / 15  & 11 / 18  \\
%java.lang.IndexOutOfBoundsException       & 47  / 15 & 9,75\%   & 127 / 39  & 3,12\% & 47 / 2  & 0/0  & 71/22  & 8 / 14    \\
%java.lang.NoSuchMethodError               & 40 / 14  & 8,30\%  & 57 / 23  & 1,40\%  & 9  /  0  & 0  / 0 & 39 /  17    & 9 / 6  \\
%java.util.ConcurrentModificationException & 38 / 5 & 7,88\% & 54 / 11  & 1,33\%   & 5  / 0  & 0  / 0 & 43 / 3 & 6 / 7  \\
%
%java.lang.NullPointerException	&	254	/	78	&	52,70\%	&	1225	/	439	&	30,08\%	&	473	/	52	&	18	/	2	&	595	/	241	&	137	/	143	\\	%LIBORAPP=408
%java.lang.IllegalStateException	&	99	/	21	&	20,54\%	&	234	/	44	&	5,75\%	&	165	/	20	&	12	/	19	&	36	/	5	&	20	/	19		\\
%java.lang.IllegalArgumentException	&	94	/	48	&	19,50\%	&	255	/	98	&	6,26\%	&	146	/	49	&	6	/	6	&	64	/	31	&	18	/	26		\\
%java.lang.RuntimeException	&	91	/	31	&	18,88\%	&	232	/	87	&	5,70\%	&	167	/	36	&	1	/	1	&	47	/	17	&	17	/	34		\\
%java.lang.OutOfMemoryError	&	56	/	22	&	11,62\%	&	180	/	57	&	4,42\%	&	121	/	20	&	15	/	1	&	17	/	18	&	23	/	11		\\
%java.lang.NoClassDefFoundError	&	52	/	15	&	10,79\%	&	73	/	21	&	1,79\%	&	9	/	1	&	0	/	0	&	37	/	9	&	26	/	11		\\
%java.lang.ClassCastException	&	49	/	15	&	10,17\%	&	94	/	36	&	2,31\%	&	43	/	12	&	0	/	0	&	40	/	15	&	11	/	9		\\
%java.lang.IndexOutOfBoundsException	&	47	/	15	&	9,75\%	&	127	/	39	&	3,12\%	&	47	/	6	&	0		0	&	71	/	22	&	8	/	10		\\
%java.lang.NoSuchMethodError	&	40	/	14	&	8,30\%	&	57	/	23	&	1,40\%	&	9	/	1	&	0	/	0	&	39	/	17	&	9	/	5		\\
%java.util.ConcurrentModificationException	&	38	/	5	&	7,88\%	&	54	/	11	&	1,33\%	&	5	/	0	&	0	/	0	&	43	/	3	&	6	/	7

@@

We can observe that most of this set is composed by exceptions defined by java.lang,
which are mostly implicitly thrown by the JVM due to programming mistakes 
(e.g., out-of-bounds array index, division-by-zero, access to a null reference)
 or resource limitation (e.g., OutOfMemoryError).
For such exceptions there is usually no proper handling besides presenting an error message to
 the user and restarting the application. Only high fault tolerant systems need to 
provide solutions to handle them.

\emph{\textbf{The high number of NullPointerExceptions.}}
From this set the java.lang.NullPointerException was the most reported root cause (27,71\% )
Considering the frequencly of NullPointerException across projects, we could observe that
 51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
was the root cause.
 The NullPointerExceptions are mainly signaled inside the applicaiton code and the Android platform,
 although we also find NullPointerExceptions being signaled from third-party libraries. Regarding
reusable code, there is no consensus whether it is a good or bad practice to 
re-throw a NullPointerException. Some prefer to encapsulate such an exception on
IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
should not be wrapped.

The high prevalence of NullPointerExceptions is aligned with the findings of other 
works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs 
related to exception handling were caused by NullPointerException; other works on robustness 
testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
detected bugs were due to NullPointerExceptions 
and exceptions implicitly-signaled by Java environment due to programming mistakes or resource limitations
 (as the ones found in this study).In Section ~\ref{sec:disc}  we provide further 
discussion on the NullPointerException problem, pointing 
to ways of how to mitigate it.

%\emph{\textbf{Direct instances of RuntimeExceptions being thrown.}}
%From Table ~\ref{tab:topten} we can observe that direct instances
%java.lang.RuntimeException were reported as the root cause of stacks in 19,09\% 
%of the projects, and most of such exceptions  were  thrown by the Android platform (167 out of 232).
%Throwing general exceptions, such as direct instances of RuntimeException, violates best
%practice III. It is considered a bad practice because exception type does not carry enough
%information to identify the cause of the exceptional behavior, and as a consequence developers need to relying on
%the exception message which can be neither complete nor precise~\cite{gosling2000java}.

\emph{\textbf{The prevalence of programming mistakes.}} To get a broader view of the prevalence of
programming mistakes of reported exception stack traces we performed the following analysis....

 we performed an empirical evaluation whose goal was to identify the concerns related to the most reported root exceptions. 
Since it is infeasible to inspect the code responsible for throwning every exception reported in this study.
This identification performed in this study was based on meaning that the name of the exception holds, 
its description (i.e., Javadoc documentaion), and the Java specification. To exemplify, an instance of ArrayOutOfBoundException
 usually refers to a programming mistake. Moreover, the Java specification list all exceptions related to backward compatibility 
issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError). Some concerns are known as sources 
of faults in mobile development are: concurrency ~\cite{ama2012} backward compatibility  
~\cite{McDon13}, security  ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
However, there are exceptions that cannot be directly mapped to one of such concerns. Either because they are
related to programming mistakes - as illustrated on the previously. Or because they are too general (i.e.,  java.lang.Exception,
java.lang.RuntimeException, java.lang.Error). Or also because they are related to other concerns (e.g., specific to a given library).

%The list of categories is illustrated in Table~\ref{tab:categories}.

To perform such mapping, we selected a subset of all reported root exceptions. This subset was composed by 100 root exceptions 
which were reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the source code and
 Javadoc related to each exception,  and Java specification, we identified the concern
releated to each root exception. Than we could calculate the prevalence of a concern on the reported stack traces using the following metric.
Given that: Stacks is the complete set of stacks found in this study, Lc is the list of root causes related to a given concern (as identified on the study),
 and Stacks[Lc] the list of issues that mentions the concern. The prevalence of a concern could be given by the formula: Stacks[Lc]/Stacks. 
Table~\ref{tab:tophundrend} illustrates the result of this empirical evaluation.

% When the root exceptions that were too general, or related to other concerns 
% were also maped to these additional categories. Table~\ref{tab:tophundrend} illustrates the result of such mapping.%
%\begin{table}
%  \centering
%  \begin{tabular}{|p{2cm}| p{5cm}|}
%    \hline
%    \bfseries{Category} & \bfseries{Description} \\
%    \hline
%      Programming logic &  Exceptions thrown by the JVM when the 
% evaluation of an expression violates the normal semantics of language (e.g., 
% out-of-bounds array index, division-by-zero, access to a null reference). \\ \hline
%      Resources (IO, Memory, Batery)  & Exception releated to the manipulation or restriction of resources (e.g., file, network, memory) \\ \hline
%      Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
%      Concurrency                            & Exceptions related to multi-threaded programming \\ \hline
%      Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
%      Specific Exceptions              & Exceptions created for specific applications, frameworks or libraries \\ \hline
%      General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
%    \hline
%  \end{tabular}
%  \caption{Description of categories.}
%  \label{tab:categories}
% \end{table}

\begin{table}
  \centering
  \begin{tabular}{lrr}
    \hline
    \bfseries{Category} &  \bfseries{\% Occurrences on stacks} \\
    \hline
      Programming logic (java.lang and util) &  52,0\%\\ 
      Resources (IO, Memory, Batery)       &   23,9\% \\ 
      Security                               &  4,1\%\\  
      Concurrency                            &  2,9\% \\ 
      Backward compatibility                 & 5,5\% \\ 
      Specific Exceptions               &  4,9\%\\ 
      General (Error, Exception, Runtime)    &  6,7\%\\
  \end{tabular}
  \caption{Identifying the concerns related to root exceptions}
  \label{tab:tophundrend}
\end{table}

To ensure the quality of the process, three independent coders classified a randomly selected
sample of 25 exception types (from the total 100) using the same list of concerns;
the interater agreement was high (96\%). We could also observe that
We can therefore assume that the categories are descriptive 
enough for covering the range of exception types analyzed.

These study showed that the programming mistakes associated to the most reported root causes. 
However, they may be interactly related to one of the other concerns. For  instance, a concurrency problem may 
lead to a NullPointerException.

@@

%TODO: Discuss about the interater agreement here...
%\subsection{Sizes of Stack traces}
%Concerning the size of the stacks (i.e. the number of frames composing them), we
%could observe that  sometimes it exceeded 100 frames (158 stack frames).
%Manually inspecting the corresponding execution trace was caused by one of the
%two reasons: recursive calls or exceptions signaled by a method defined deep
%down in a reused framework; exceptions successively wrapped. Irrespective of the
%reason, in exception flows involving many methods, it is practically impossible
%to handle an exception in a way other than exiting the application, as at the
%place the exception occurs there is hardly enough contextual information to
%perform recovery actions.

%Considering only the packages of each exception we could observe that 81,28\%
%of all analyzed set of stack traces reported exceptions defined by java.lang
%(see Table XXX).

%\begin{table}
% \centering
%\begin{tabular}{lrrrr}
%    \hline
% \bfseries{rank} & \bfseries{package} & \bfseries{occurr} &  \bfseries{\%occur} & % \bfseries{projects} \\
%     \hline
%1 & java.lang         & 19471 & 81,28\% & 3080 \\
%2 & java.io           & 1222  & 5,10\%  & 509 \\
%3 & java.net          & 740   & 3,09\%  & 311 \\
%4 & java.util         & 689   & 2,88\%  & 287 \\
%5 & java.lang.reflect & 115   & 0,48\%  & 78 \\
%\hline
%  \end{tabular}
%\caption{Top 5 most popular packages associated to the root exceptions of stack traces}
%\label{tab:toptenpopular2}
%\end{table}

%\noindent \fbox{
%\begin{minipage}{0.96\columnwidth} 
%\emph{RQ2: Most exceptions are due to programming mistakes
%or generally irrecoverable errors.}
%\end{minipage}}

@@

 Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on the analyzed issues. As we can see 
from Table ~\ref{tab:typeroottab},  most of the reported exceptions signaled (explicilty or implicitly by the JVM) on methods defined
 on the Android Plaform on on the application itself.

@@

\emph{\textbf{Undocumented Runtime Exceptions}}.
However, we can also observe a considetable amount of 
runtime exception signaled on library reused by apps.
In order to evaluate if  best practice IV was being obeyed, we inspected the   
signature of each method signaling a runtime exception. 
We inspected methods of the Android Plaftorm and 
the libraries, because both are  third  party code reused 
by the Android apps. In this task we filtered out all the exceptions implicitly 
signaled by  JVM (due to programming mistakes) 
since these exceptions should not be documented on the 
method signature.

@@

This analysis revealed unexpected wrappings such as: Error wrapping a Checked 
exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.
Java is the only language that provides a hybrid exception model 
which offers  different types to represent different exception behaviors (i.e., error, 
runtime and checked) (see Section~\ref{sec:best}). For instance,
according to Java specification, Errors should not be handled inside the system 
since they usually represent unrecoverable conditions
detected by the JVM (e.g.,  OutOfMemoryError). On the other hand, checked exceptions should
represent recoverable conditions. Next we discuss and exemplify the cross-type wrappings
 found in this study.

@@

Hence, when (mis)applied, the exception wrapping can make the exception-related code
 more complex and lead to what we call an \emph{exception handling confusion problem}.
Such problem can lead the program to
an unpredictable state in the presence of exceptions. To illustrate this problem, we can use
 one of the examples found in this study: when the developer is 
confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition, however we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the JVM which make impossible the program to recover.

@@

Besides the explicitly thrown runtime exceptions, in this work we also observe a high number of 
stack traces reporting coding errors (out-of-bounds array index, division-by-zero, access to a null).
Such exceptions are implicitly thrown by JVM, and among them the NullPointerException 
were the reported root cause in most of the analyzed exception stack traces.

@@

\textit{Analysis and Use of Stack Trace Information.} Several works have
investigated the use of stack trace information to support bug classification
and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},
fault-proneness prediction models~\cite{kim2013predicting} and even automated
bug fixing tools~\cite{sinha2009fault}. Kim et al.~\cite{kim2011crash} use an
aggregated form of multiple stack traces available in crash reports to detect
duplicate crash reports and to predict if a given crash will be fixed. Dhaliwal
et al.~\cite{dhaliwal2011classifying} proposed a crash grouping approach that
can reduce bug fixing time in approximately 5\%. Wang et
al.~\cite{wang2013improving} propose an approach to identify correlated crash
types and describe a fault localization method to locate and rank files related
to the bug described on a stack trace. Schroter et al.~\cite{schroter2010stack}
conducted an empirical study on the usefulness of stack traces for bug fixing
and showed that developers fixed the bugs faster when failing stack traces were
included on bug issues.  In a similar study, Bettenburg et
al.~\cite{bettenburg2008makes} identify stack traces as the second most stack
trace feature for developers.  Sinha et al.~\cite{sinha2009fault} proposed an
approach that uses stack traces to guide a dataflow analysis for locating and
repairing faults that are caused by the JVM implicitly signaled exceptions. Kim
at al.~\cite{kim2013predicting} proposed an approach to predict the
crash-proneness of methods based information extracted from stack traces and
methods' bytecode operations.  They observed that most of the stack traces were
related to NullPointerException and other JMV implicitly thrown exceptions had
the higher prevalence on the analyzed set of stacks.

>>>>

Besides, the runtime exceptions are also implicitly thrown by the runtime environment when a program violates 
the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
error, null pointer references).\footnote{Some programming languages react to such errors by peremptorily terminating the program; 
other programming languages allow similar approaches e.g.C\#; and others let the program to continue
 its execution in some situations such as the out-of-bounds array index e.g., C++. }  
The programs are not expected to handle such runtime exceptions signaled by the runtime environment ~\cite{gosling2000java}.

@@

In Java, once an exception is thrown, the runtime environment looks for the nearest enclosing exception handler
(in Java it is represented by the try-catch block), and unwinds the execution stack if necessary.
 Looking for the handler on the invocation stack is claimed to increase software reusability, 
since the invoker of an operation can handle it in a wider context ~/cite{miller1997issues}.

@@

\emph{II-Error represents an unrecoverable condition detected by the runtime enviroment which
should not he handled} (\cite{gosling2000java}). It results from failures detected
by the Java Virtual Machine which indicate resource deficiencies, invariant
failures or other conditions that make impossible the program to recover.

@@

\emph{III-A method should throw exceptions that precisely define the
exceptional condition} (\cite{gosling2000java,bloch2008effective}). To do so,
developers should either try to reuse the exception types already defined in the
Java API or they should create a specific exception. Throwing general types such as a
pure java.lang.Exception or a java.lang.RuntimeException is considered a bad practice.

@@

\emph{IV- All exceptions explicitly thrown by a method should be documented.}
(\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled runtime exceptions related to programming
mistakes} as carefully as checked exceptions. Doing so, the clients of a method
will be aware of all the exceptions the method can throw and can protect design the code
to deal with them ~\cite{Robil00}.

@@

@@

\subsection{Manual Inspections}
In the context of the ExceptionMiner tool, manual inspections were used:
to help on the identification of packages composing the Android platform, 
libs and apps analyzed in this study (as described previously); and to 
identify the type of some exceptions reported on issues, 
that could not be automatically identified by the ExceptionMiner tool.
Some exceptions could not be automaticaly identified because they were 
not defined on the last version of the application (the one analyzed by ExceptionMiner),
or were defined on libraries not found among application resources.
In such cases we manually inspect the sourcecode or javadoc
 documentation available online of each of them. 
When the exception was not found on the previous versions, the exception where classified 
in our study as "Undefined".  Only 31 exceptions 
remained undefined - such exceptions were reported on 61 exception stack traces as 
presented in Table ~\ref{tab:typeroottab}.

@@

This section presents the results of the study, providing both a
quantitative and qualitative analysis of the outcomes. The structure of this section
 is centered on questions that guied the search for fragilities (on the exception-related code)
based on exception stack trace information.

@@

After distilling the information available on the exception stack traces we could find 
the the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:topten} presents a list the top 10 root exceptions found in the study - 
 ranked by the number of distinct projects on which they were reported. 
The table shows how many times the signaler of such exception was a method defined on
the Android Platform, the Android libcore, the application itself or a third-party library -
 following the classification as presented in Table  \ref{tab:signalers}.

@@

We can observe that most of the exceptions on this list are implicitly thrown by the
 runtime environment due to programming mistakes  (e.g., out-of-bounds array index, division-by-zero, access to a null reference)
 or resource limitation (e.g., OutOfMemoryError). 
From this set the java.lang.NullPointerException was the most reported root cause (27,71\% ). 
If we consider the frequencly of NullPointerException 
across projects, we can observe that 51,96\% of all projects reported at least one exception stack 
trace on which the NullPoiterException was the root cause.
 The NullPointerExceptions are mainly signaled inside the applicaiton code and the Android platform,
 although we also find NullPointerExceptions being signaled from third-party libraries. Regarding
reusable code, there is no consensus whether it is a good or bad practice to 
re-throw a NullPointerException. Some prefer to encapsulate such an exception on
IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
should not be wrapped.

The high prevalence of NullPointerExceptions is aligned with the findings of other 
works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs 
related to exception handling were caused by NullPointerException; other works on robustness 
testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
detected bugs were due to NullPointerExceptions and exceptions implicitly-signaled by Java 
environment due to programming mistakes or resource limitations
 (as the ones found in this study).

\emph{\textbf{The prevalence of programming mistakes.}} To get a broader view of the prevalence of
programming mistakes of reported exception stack traces we: (i) identified the Java exceptions that can be
implicitly thrown by the runtime environment due to a programing mistake or resource limitation; and (ii)
 used ExceptionMiner to evaluate the prevalence of such exceptions on the reported stack traces.
Table ~\ref{tab:causes} illustrates the main causes of exceptions reported on issues.

\begin{table}
  \centering
  \begin{tabular}{lrr}
    \hline
    \bfseries{Cause} &  \bfseries{\% Occurrences on stacks} \\
    \hline
      Programming logic (java.lang and util) &  52,0\%\\ 
      Resources Restriction   &   23,9\% \\ 
      General Exceptions (Error, Exception, Runtime)    &  6,7\%\\
      Specific  &  xxx\%\\ 
 \end{tabular}
  \caption{Identifying the concerns related to root exceptions}
  \label{tab:causes}
\end{table}

The analysis on the complete set os stack traces revealed that, in fact, the programming mistakes were
the most reported root causes.  Although there is usually no proper handling for such exceptions, 
besides presenting an error message to the user and restarting the application ~\footnote{Only 
high fault tolerant systems need to provide solutions to handle them.}.

This finding calls the attention to the fact that: high number of programming errors reported 
on issues may me an indication that developers are underestimating the impact that such 
mistakes have on application robustness.

@@

@@

 Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on the analyzed issues. As we can see 
from Table ~\ref{tab:typeroottab},  most of the reported exceptions signaled (explicilty or implicitly by the runtime enviroment) on methods defined
 on the Android Plaform on on the application itself.

@@

\emph{\textbf{Undocumented Runtime Exceptions}}.
However, we can also observe a considetable amount of 
runtime exception signaled on library reused by apps.
In order to evaluate if  best practice IV was being obeyed, we inspected the   
signature of each method signaling a runtime exception. 
We inspected methods of the Android Plaftorm and 
the libraries, because both are  third  party code reused 
by the Android apps. In this task we filtered out all the exceptions implicitly 
signaled by  runtime environment (due to programming mistakes) 
since these exceptions should not be documented on the 
method signature.

@@

This analysis revealed unexpected wrappings such as: Error wrapping a Checked 
exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.
Java is the only language that provides a hybrid exception model 
which offers  different types to represent different exception behaviors (i.e., error, 
runtime and checked) (see Section~\ref{sec:best}). For instance,
according to Java specification, Errors should not be handled inside the system 
since they usually represent unrecoverable conditions
detected by the runtime environment (e.g.,  OutOfMemoryError). On the other hand, checked exceptions should
represent recoverable conditions. Next we discuss and exemplify the cross-type wrappings
 found in this study.

@@

Hence, when (mis)applied, the exception wrapping can make the exception-related code
 more complex and lead to what we call an \emph{exception handling confusion problem}.
Such problem can lead the program to
an unpredictable state in the presence of exceptions. To illustrate this problem, we can use
 one of the examples found in this study: when the developer is 
confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition, however we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the runtime environment which make impossible the program to recover.

@@

Besides the explicitly thrown runtime exceptions, in this work we also observe a high number of 
stack traces reporting coding errors (out-of-bounds array index, division-by-zero, access to a null).
Such exceptions are implicitly thrown by runtime environment, and among them the NullPointerException 
were the reported root cause in most of the analyzed exception stack traces.

@@

\textit{Analysis and Use of Stack Trace Information.} Several works have
investigated the use of stack trace information to support bug classification
and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},
fault-proneness prediction models~\cite{kim2013predicting} and even automated
bug fixing tools~\cite{sinha2009fault}. Kim et al.~\cite{kim2011crash} use an
aggregated form of multiple stack traces available in crash reports to detect
duplicate crash reports and to predict if a given crash will be fixed. Dhaliwal
et al.~\cite{dhaliwal2011classifying} proposed a crash grouping approach that
can reduce bug fixing time in approximately 5\%. Wang et
al.~\cite{wang2013improving} propose an approach to identify correlated crash
types and describe a fault localization method to locate and rank files related
to the bug described on a stack trace. Schroter et al.~\cite{schroter2010stack}
conducted an empirical study on the usefulness of stack traces for bug fixing
and showed that developers fixed the bugs faster when failing stack traces were
included on bug issues.  In a similar study, Bettenburg et
al.~\cite{bettenburg2008makes} identify stack traces as the second most stack
trace feature for developers.  Sinha et al.~\cite{sinha2009fault} proposed an
approach that uses stack traces to guide a dataflow analysis for locating and
repairing faults that are caused by the implicitly signaled exceptions. Kim
at al.~\cite{kim2013predicting} proposed an approach to predict the
crash-proneness of methods based information extracted from stack traces and
methods' bytecode operations.  They observed that most of the stack traces were
related to NullPointerException and other JMV implicitly thrown exceptions had
the higher prevalence on the analyzed set of stacks.
